---
title: "The Great American Coffee Taste Test Analysis"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{The Great American Coffee Taste Test Data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This dataset, provided by [TidyTuesday project](https://github.com/rfordatascience/tidytuesday/blob/master/data/2024/2024-05-14/readme.md), contains 4042 observations and 57 variables related to coffee consumption preferences in the U.S.

## Libraries

To analyze the data and create visualizations, you'll need the following libraries:

```{r , warning=FALSE,eval=FALSE, echo=TRUE, collapse=FALSE}
# Install the necessary packages if not already installed
install.packages(c("dplyr", "ggplot2", "plotly", "DT"))

```

```{r ,  warning=FALSE,collapse=FALSE, message=FALSE}
# Load the libraries
library(CoffeeSurvey)
library(dplyr)
library(ggplot2)
library(plotly)
library(DT)
```


I have filtered the dataset to focus on 6 key variables for my analysis. If you are interested in exploring the full dataset with 57 variables, you can download it directly from the TidyTuesday GitHub repository [here](https://github.com/rfordatascience/tidytuesday/blob/master/data/2024/2024-05-14/readme.md).

The data can be accessed by loading "coffee_survey" or alternatively using the code "CoffeeSurvey::coffee_survey".

You can use the following code to have a look on the dataset.

```{r}
data("coffee_survey")
head(coffee_survey)
```

## Analysis Examples

Below are two simple examples showing how to explore the data through a table and a plot. Feel free to further explore using similar methods.

### 1. Coffee Consumption by Location

This first example shows where respondents typically drink coffee (home, office, cafe, etc.).

```{r}
# Filter and calculate percentages for coffee consumption locations
where_drink_distribution <- coffee_survey |>
  filter(!is.na(where_drink)) |>
  filter(where_drink %in% c("At home", "At the office", "At a cafe", "On the go")) |>
  count(where_drink) |>
  mutate(percentage = n / sum(n) * 100) |>
  arrange(desc(percentage))
```

```{r table1}
#| label: tbl-where
#| tbl-cap: "Where do you typically drink coffee?"

summary_table <- where_drink_distribution |>
  select(
    Location = where_drink,                  
    `Percentage Responded` = percentage        
  ) |>
  mutate(`Percentage Responded` = round(`Percentage Responded`, 1)) |>  
  arrange(desc(`Percentage Responded`))

datatable(summary_table)
```

The majority of respondents (86.9%) reported consuming coffee at home, followed by 7.5% at the office, 3.8% at a cafe, and 1.7% while on the go, highlighting a strong preference for home-based coffee consumption.

### 2. Coffee Type Preferences

This second example visualizes the overall preferences for different coffee types.

```{r}
# Define colors for coffees
coffee_colors <- c(
  "Coffee A" = "#8B4513",
  "Coffee B" = "#D2B48C",
  "Coffee C" = "#C19A6B",
  "Coffee D" = "#4B3832")

popularity <- coffee_survey |>
  filter(!is.na(prefer_overall)) |>
  count(prefer_overall) |>
  mutate(percentage = n / sum(n) * 100) |>
  arrange(desc(n))

```

```{r figure 1}
#| label: fig-bar1
#| fig-cap: "Overall Perferences"

gg1 <- ggplot(popularity, aes(x = prefer_overall, y = percentage, fill = prefer_overall)) +
      geom_bar(stat = "identity") +
      labs(title = "Popularity of Coffee Types",
           x = "Coffee Type",
           y = "Percentage of Votes") +
      scale_fill_manual(values = coffee_colors) +
      theme_minimal() +
      theme(
        plot.title = element_text(size = 16, face = "bold", hjust = 0.5),
        axis.title.x = element_text(size = 12),
        axis.title.y = element_text(size = 12)
      )

ggplotly(gg1)
```

Coffee D (the experimental Colombian coffee) was the most popular choice, with 36.7% of the votes. This suggests that consumers are open to trying more unusual, fermented flavor profiles. Coffee A (light-roast Kenyan) followed closely behind with 21.7% of the votes, indicating a strong market for lighter, fruitier coffee options.

## Explore Further

Feel free to explore other variables, such as brewing methods, price sensitivity, and demographic trends. If you're interested in the full dataset, you can download it from the TidyTuesday GitHub repository linked above.
